---
title: "NYPD Shooting Incident Data Analysis"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Input file**

* Title: NYPD Shooting Incident Data (Historic)
* Url: [https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv](https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv)
* Dataset description:  List of every shooting incident that occurred in NYC going back to 2006 through the end of the previous calendar year.



```{r import_data}
options(repr.plot.width=30, repr.plot.height=8)
library(tidyverse)

url_in <- 'https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv'
NYPD <- read_csv(url_in)
```


## Check the data structure
```{r check_stucture}

library(dplyr)
glimpse(NYPD)  
```


## Possible bias
1. Different boroughs may have different security levels, which means they have different numbers of shooting incidents. Brooklyn probably has a higher crime rate than other boro's.
2. Different age groups can have different shootings incident rate. 20s may be more inclined to shoot.


## Analytics plan
1. Will check the number of incident by boro and age group to verify the bias above.
2. Also leverage modeling method to find the relation between number of death and number of incidents.




## Data transform : Change OCCUR_DATE to date format
```{r change_date}

NYPD$OCCUR_DATE <- as.Date(NYPD$OCCUR_DATE,format='%m/%d/%Y')

```


## Count the number of incident by each boro in NY in 2020 to see which area has more cases
```{r summarize_boro}
NYPD_by_boro_2020 <- NYPD %>% 
  mutate(OCCUR_DATE = as.Date(OCCUR_DATE,format='%m/%d/%Y') ) %>% 
  filter( between(OCCUR_DATE, as.Date("2020-01-01"), as.Date("2020-12-31"))  ) %>% 
  group_by(BORO) %>%
  summarize(cases =n()) %>%
  select(BORO,cases) %>%
  ungroup()

NYPD_by_boro_2020
```

## Get the boro with the highest number of shooting incident

```{r top_boro}
NYPD_by_boro_2020 %>% 
  slice_max(cases, n=1)
```

## Get the death rate of shooting incident for each boro in 2020
```{r summarize_death_boro}
NYPD_death_rate_by_boro <- NYPD %>%
    mutate(OCCUR_DATE = as.Date(OCCUR_DATE,format='%m/%d/%Y') ) %>% 
    filter( between(OCCUR_DATE, as.Date("2020-01-01"), as.Date("2020-12-31"))  ) %>% 
    group_by(BORO) %>%
    summarize(cases =n(),deaths = sum(STATISTICAL_MURDER_FLAG)) %>%
    mutate(deaths_rate = round(deaths / cases,3)) %>%
    ungroup()
NYPD_death_rate_by_boro
```

## Get the boro with the highest death rate in shooting incident

```{r top_death_rate_boro}
NYPD_death_rate_by_boro %>% 
  slice_max(deaths_rate, n=1)
```




## Count the number of incident by Perpetrator’s age group in 2020
```{r summarize_age}
NYPD_by_age <- NYPD %>%
    mutate(OCCUR_DATE = as.Date(OCCUR_DATE,format='%m/%d/%Y') ) %>% 
    filter( between(OCCUR_DATE, as.Date("2020-01-01"), as.Date("2020-12-31"))  ) %>% 
    filter(! is.na(PERP_AGE_GROUP)  ) %>%
    group_by(PERP_AGE_GROUP) %>%
    summarize(cases =n()) %>%
    select(PERP_AGE_GROUP,cases) %>%
    ungroup()
NYPD_by_age
```




## Get the age group with the highest number of incident
```{r top_age}
NYPD_by_age %>% 
  slice_max(cases, n=1)
```


## Visualize number of cases by boro 
```{r chart_boro}
options(repr.plot.width=30, repr.plot.height=8)
ggplot(NYPD_by_boro_2020, aes(reorder(BORO, cases, sum), cases)) +  geom_col(fill = "#4472C4") +
  geom_text(aes(label=cases), position=position_dodge(width=0.9), hjust=0) + 
  coord_flip() +   
  labs(title = "Number of Shooting Incidents by Borough in 2020", y= NULL)+
  theme(text=element_text(size=10)) 
```

## Visualize death rate per incident by boro 
```{r chart_date_rate_boro}
options(repr.plot.width=30, repr.plot.height=8)
ggplot(NYPD_death_rate_by_boro, aes(reorder(BORO, deaths_rate, sum), deaths_rate)) +  geom_col(fill = "#ED7D31") +
  geom_text(aes(label=deaths_rate), position=position_dodge(width=0.9), hjust=-0.1) + 
  coord_flip() +   
  labs(title = "Death Rate per Shooting Incident by Borough in 2020", y= NULL) + 
  theme(text=element_text(size=10)) #change font size of legend title
```


## Visualize number of cases by age group
```{r chart_age}
options(repr.plot.width=30, repr.plot.height=8)
ggplot(NYPD_by_age, aes(reorder(PERP_AGE_GROUP, cases, sum), cases)) +  geom_col(fill = "#00BFC4") + coord_flip() +
    geom_text(aes(label=cases), position=position_dodge(width=0.9), hjust=-0.1) + 
    labs(title = "Number of Shooting Incidents by Age", y= NULL) +
    theme(text=element_text(size=10))
```


## Build model to see the relationship between number of deaths and number of shooting incidents
```{r prep_for_model}

NYPD_by_month <- NYPD %>%
    mutate(OCCUR_DATE = as.Date(OCCUR_DATE,format='%m/%d/%Y') ) %>%
    mutate(OCCUR_MONTH = strftime(OCCUR_DATE,format='%Y/%m') ) %>%
    mutate(Month = strftime(OCCUR_DATE,format='%m') ) %>%
#     filter( between(OCCUR_DATE, as.Date("2020-01-01"), as.Date("2020-12-31"))  ) %>% 
    group_by(OCCUR_MONTH,Month) %>%
    summarize(deaths = sum(STATISTICAL_MURDER_FLAG), cases= n())

```


## Build model to see the relationship between number of deaths and number of shooting incidents

```{r model}
mod <- lm(deaths  ~ cases  , data = NYPD_by_month)
summary(mod)
```


## Predict number of deaths with model

```{r pred_death}
pred <- tibble(pred = predict(mod))

NYPD_by_month_w_pred <- cbind(NYPD_by_month,pred)
```


## Plot predicted deaths and actual deaths

```{r plot_death}
options(repr.plot.width=30, repr.plot.height=8)
NYPD_by_month_w_pred %>% ggplot() +
   geom_point(aes(x=cases, y=deaths, size = 1.5), color = "blue") +
   geom_point(aes(x = cases, y = pred,size = 1.5), color = "red") +
   geom_line(aes(x = cases, y = pred ),linetype = "dashed", color = "red") +
   labs(title = "Model deaths with cases", y= NULL) +
   theme(
#        legend.position="bottom",
    text=element_text(size=10)) #change font size of legend title
```



## Build model to see the relationship between number of shooting incidents and calendar month

```{r model2}
mod_m <- lm(cases ~ Month , data = NYPD_by_month)
summary(mod_m)
```

## Predict number of cases with model

```{r pred_cases}
pred_m <- tibble(pred_m = predict(mod_m))

NYPD_by_month_w_pred <- cbind(NYPD_by_month_w_pred,pred_m)
```


## Plot predicted deaths and actual deaths

```{r plot_case_model}
options(repr.plot.width=30, repr.plot.height=8)
NYPD_by_month_w_pred %>% ggplot() +
   geom_point(aes(x=Month, y=cases, size = 0.5), color = "blue") +
   geom_line(aes(x = Month, y = pred_m ),linetype = "dashed", color = "red") + 
   geom_point(aes(x = Month, y = pred_m,size = 1.5), color = "red") +
   labs(title = "Model cases with month", y= NULL) +
   theme(
#        legend.position="bottom",
    text=element_text(size=10)) #change font size of legend title

```



## Conclusion
1. Brooklyn has had more shootings than any other borough.
2. The Staten Island has the highest fatality rate from shootings.
3. In 2020 in NY, the shootings were mainly committed by people who are 25~44.
4. 18% chance of dying in a shooting incident.
5. There were more shootings in July and August than any other month.








##	my session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```












